Restrictive Clearance Comparison

### 2025-03-19 data-raw/ClearanceProject-2-RelativeLikelihood.Rmd
Merge feature/Rclearance_comparison to dev (#22)
* Added diff_log_cvt() call for oral and iv.

* Changed log_likelihood() to work with pre-calculated predictions from other sources. Still requires sigma.

* Quick fix to get_fit.pk which no longer needs unnesting.

* Renamed and initialized 4 analysis documents for clearance project.

* Added data associated with clearance project for efficient loading.

* Organized ongoing hepatic clearance project files in data-raw/

@@

```{r venn_diagrams}
member_list <- list(
  httk_3comp2 = unique(get_cheminfo(info = "DTXSID",
                                species = "human",
                                model = "3compartment2",
                                median.only = TRUE,
                                physchem.exclude = TRUE,
                                suppress.messages = TRUE)) %>% 
    suppressWarnings(),
  cvt = cvt %>% 
    filter(species %in% c("human", "rat"),
           conc_medium_normalized == "plasma") %>%
    pull(analyte_dtxsid)
)

ggvenn(member_list, fill_color = c("white", "white"))

@@

target_chems <- intersect(member_list$cvt, member_list$httk_3comp2)

filter_targets <- function(.data) {
  stopifnot("Chemical" %in% names(.data))
  dplyr::filter(.data = .data, Chemical %in% target_chems)
}
```

@@

```{r selecting_chemical_params}
chosen_list <- rat_pars_human
# The following are derived from the above selected list (see environment)
this_chem_v <- unique(chosen_list$param_df$Chemical)
this_param_rest_l <- chosen_list$named_params_restrictive
this_param_nrest_l <- chosen_list$named_params_nonrestrictive
this_param_df <- chosen_list$param_df
this_httk_clearance_df <- chosen_list$clearance_df
this_species_v <- c("rat", "human")
# Testing to make sure there are some chemicals that have different clearances
count(this_httk_clearance_df, restrictive_clearance == nonrestrictive_clearance)
# The majority (140) do!
```

Here I do a quick fit of the CvT data only for those that have plasma measurements.

```{r quick-fit}

# Only use plasma for simplicity
this_pk <- pk(cvt %>% 
                dplyr::filter(conc_medium_normalized == "plasma")
              ) +
  settings_optimx(method = "bobyqa") +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(
    keep_data_original = TRUE,
    suppress.messages = FALSE) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  stat_error_model(error_group = vars(Chemical, Species))

this_pk <- do_prefit(this_pk)
this_pk <- do_fit(this_pk, n_cores = 14)
```

```{r load-other}
# Here I want to load all other data and setup other data objects
# Kd/binding affinity data
load(paste0(Sys.getenv("OD_DIR"),
            "Profile/Documents/New-HTTK-raw-2019-05-06.RData")
)

# Chemical info and properties (averaging the latter by chemical and experimental/predicted)
target_chem_info <- get_chemical_details_batch(DTXSID = target_chems,
                                               rate_limit = 1L)
target_chem_properties <- get_chem_info_batch(DTXSID = target_chems, 
                                              rate_limit = 1L) %>%
  filter(name %in% c("LogKow: Octanol-Water",
                     "pKa Acidic Apparent",
                     "pKa Basic Apparent")) %>% 
  group_by(dtxsid, propType, propertyId) %>%
  summarize(Mean = mean(value, na.rm = TRUE),
            StDev = sd(value, na.rm = TRUE)) %>%
  ungroup() %>%
  distinct()

# Bring together all chemical properties
target_chem_pw_predicted <- target_chem_properties %>%
  filter(propType == "predicted") %>%
  pivot_wider(names_from = propertyId,
              names_prefix = "pr_",
              id_cols = dtxsid,
              values_from = c("Mean", "StDev"))
target_chem_pw_experimental <- target_chem_properties %>%
  filter(propType == "experimental") %>%
  pivot_wider(names_from = propertyId,
              names_prefix = "ex_",
              values_from = c("Mean", "StDev"),
              id_cols = dtxsid)

target_chem_pw_full <- full_join(
  target_chem_pw_experimental,
  target_chem_pw_predicted
) %>%
  janitor::clean_names() %>%
  mutate(dtxsid,
         type_logkow = ifelse(is.na(mean_ex_logkow_octanol_water),
                              NA, "experimental"),
         logkow = ifelse(is.na(mean_ex_logkow_octanol_water),
                         mean_pr_logkow_octanol_water,
                         mean_ex_logkow_octanol_water),
         type_apKa = ifelse(is.na(mean_ex_pka_acidic_apparent),
                            NA, "experimental"),
         apKa = ifelse(is.na(mean_ex_pka_acidic_apparent),
                       mean_pr_pka_acidic_apparent,
                       mean_ex_pka_acidic_apparent),
         type_bpKa = ifelse(is.na(mean_ex_pka_basic_apparent),
                            NA, "experimental"),
         bpKa = ifelse(is.na(mean_ex_pka_basic_apparent),
                       mean_pr_pka_basic_apparent,
                       mean_ex_pka_basic_apparent),
         .keep = "unused") %>%
  mutate(type_logkow = ifelse(!is.na(logkow) & is.na(type_logkow),
                              "predicted",
                              ifelse(type_logkow == "experimental",
                                     "experimental", NA)),
         type_apKa = ifelse(!is.na(apKa) & is.na(type_apKa),
                            "predicted",
                            ifelse(type_apKa == "experimental",
                                     "experimental", NA)),
         type_bpKa = ifelse(!is.na(bpKa) & is.na(type_bpKa),
                            "predicted",
                            ifelse(type_bpKa == "experimental",
                                     "experimental", NA))) %>%
  mutate(logkow_sd = ifelse(type_logkow == "experimental",
                            st_dev_ex_logkow_octanol_water,
                            st_dev_pr_logkow_octanol_water),
         apKa_sd = ifelse(type_apKa == "experimental",
                          st_dev_ex_pka_acidic_apparent,
                          st_dev_pr_pka_acidic_apparent),
         bpKa_sd = ifelse(type_bpKa == "experimental",
                          st_dev_ex_pka_basic_apparent,
                          st_dev_pr_pka_basic_apparent)) %>%
  select(-starts_with("st_dev")) %>%
  distinct()
# Remove the parts of this final dataset
rm(target_chem_pw_experimental, target_chem_pw_predicted)

# Predictions from httk(code retained but not run after 2.20.2025)
restrictive_preds <- get_httk_preds(
  parameters = this_param_rest_l,
  pk_obj = this_pk,
  species = c("rat", "human")
) %>% rename_with(
  .fn = ~paste0(.x, "_restrictive"),
  .cols = starts_with("httk_")
)

nonrestrictive_preds <- get_httk_preds(
  parameters = this_param_nrest_l,
  pk_obj = this_pk,
  species = c("rat", "human")
) %>% rename_with(
  .fn = ~paste0(.x, "_nonrestrictive"),
  .cols = starts_with("httk_")
)

names(restrictive_preds)

all_httk_preds <- inner_join(restrictive_preds, nonrestrictive_preds) %>%
  filter_targets()
rm(restrictive_preds, nonrestrictive_preds)
# 
readr::write_csv(all_httk_preds,
                 file = paste0(Sys.getenv("FIG_DIR"),
                               "RestrictiveClearance_Analysis/",
                               "CURRENT_all_httk_preds.csv"))

all_httk_preds <- readr::read_csv(file = paste0(Sys.getenv("FIG_DIR"), 
                                                "RestrictiveClearance_Analysis/",
                                                "CURRENT_all_httk_preds.csv"))

this_nonflat_winmodel <- get_winning_model(this_pk) %>%
  filter(model %in% c("model_1comp", "model_2comp")) %>%
  filter_targets()

# Get the winning model (including flats) and only include chemicals in target_chems
this_winmodel <- get_winning_model(this_pk) %>%
    filter_targets()

winning_preds <- predict(this_pk) %>%
  semi_join(this_nonflat_winmodel) %>% # The targets already filtered here
  distinct(Chemical, Species, Chemical_Name,
           model, method, Dose, Media, Route, Reference, pLOQ, Time,
           N_Subjects, Conc, Conc_est)

# TKstats for 1-/2-compartment model
this_tkstats <- eval_tkstats(this_pk,
                             model = c("model_1comp", "model_2comp"),
                             finite_only = TRUE,
                             dose_norm = TRUE
) %>%
  filter_targets()
colnames(this_tkstats) <- gsub("/", "_", colnames(this_tkstats))
this_tkstats <- this_tkstats %>%
  inner_join(this_winmodel)

@@

human_chems <- httk::get_cheminfo(info = "DTXSID",
                                species = "human",
                                model = "3compartment2",
                                median.only = TRUE,
                                physchem.exclude = TRUE,
                                suppress.messages = TRUE)
human_chems <- intersect(human_chems, new.raw.httk.data$DTXSID)

all_hum_rat_fups <- chem.physical_and_invitro.data %>%
  filter(DTXSID %in% human_chems) %>%
  select(DTXSID, Human.Funbound.plasma, Rat.Funbound.plasma) %>%
  mutate(Rat.Funbound.plasma = as.numeric(str_extract(Rat.Funbound.plasma,
                                                      "^([^,]+)")),
         Human.Funbound.plasma = as.numeric(str_extract(Human.Funbound.plasma,
                                                        "^([^,]+)"))
         )

# Get restrictive and non-restrictive parameters
human_chems_fup <- data.frame(
  Chemical = human_chems,
  fup_restrictive = chem.physical_and_invitro.data$Human.Funbound.plasma[],
                              fup_nonrestrictive = params_nr)
```

@@

kd_v_fup_plot <- new.raw.httk.data %>%
  inner_join(all_hum_rat_fups) %>%
  pivot_longer(cols = Human.Funbound.plasma:Rat.Funbound.plasma,
               names_to = "Species", values_to = "httk_fup_med") %>%
  ggplot(aes(x = Affinity.Kd.Med, y = httk_fup_med, color = Species)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("green4", "purple2"),
                     labels = c("Human", "Rat")) +
  scale_x_log10(labels = scales::label_log(digits = 2)) +
  labs(x = "Binding affinity ratio", y = "Fraction unbound\nin plasma") +
  theme_bw() +
  theme(legend.position = "top",
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10))

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Kd_estimates_vs_fup_multispecies.png"),
       bg = "white",
       height = 3, width = 5
)

@@

## Lipinski's Rules

Here I will look at the whether chemicals are "drug-like" by evaluating if:

1. Molecular weight less than 500 Da.  
2. logP (octanol-water partitioning coefficient) no greater than 5.  
3. Less than 5 hydrogen bond donors and less than 10 acceptors.

```{r lipinksi_test_data}

target_chem_properties %>%
  ggplot(aes(x = propType, y = Mean)) +
  geom_violin() +
  geom_line(aes(group = dtxsid), color = "grey10", alpha = 0.1) +
  geom_hline(yintercept = 5, linetype = "longdash") +
  facet_grid(scales = "free_x",
             rows = vars(propertyId)) +
  scale_y_log10() +
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, angle = 45))

target_chem_properties %>%
  count(source)

target_chem_info %>%
  ggplot(aes(x = monoisotopicMass)) +
  geom_histogram(binwidth = 50, fill = "grey5",
                 boundary = 0) +
  geom_vline(xintercept = 500, linetype = "longdash") +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 100)) +
  theme_bw()

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "target_chems_molmass.png"),
       bg = "white",
       height = 3, width = 5
)

target_chem_info %>%
  count(monoisotopicMass <= 500)

target_chem_pw_full %>%
  ggplot(aes(x = forcats::fct_reorder(dtxsid, logkow))) +
  geom_pointrange(aes(y = logkow, ymin = logkow - logkow_sd,
                      ymax = logkow + logkow_sd), size = 0.1) +
  facet_grid(cols = vars(type_logkow)) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
target_chem_pw_full %>%
  ggplot(aes(x = logkow)) +
  geom_histogram(color = "grey50", fill = "grey5",
                 binwidth = 1) +
  theme_bw()

```{r reticulate-prep}
smiles_list <- target_chem_info %>% pull(smiles) %>% unique()
mykit <- import("rdkit")
```

```{python get_H_bonds}
import rdkit
from rdkit import Chem
from rdkit.Chem import Descriptors
from rdkit.Chem import rdMolDescriptors

py_smiles_list = r.smiles_list
mol_from_smiles = [Chem.MolFromSmiles(x) for x in py_smiles_list]

py_h_donors = [Chem.rdMolDescriptors.CalcNumLipinskiHBD(x) for x in mol_from_smiles]
py_h_acceptors = [Chem.rdMolDescriptors.CalcNumLipinskiHBA(x) for x in mol_from_smiles]
py_tpsa = [Chem.rdMolDescriptors.CalcTPSA(x, False, True) for x in mol_from_smiles]
py_rotbonds = [Chem.rdMolDescriptors.CalcNumRotatableBonds(x) for x in mol_from_smiles]
```

```{r lipinksi_alluvial}

target_chem_summary <- target_chem_info %>%
  distinct(dtxsid, monoisotopicMass, smiles) %>%
  right_join(data.frame(smiles = smiles_list)) %>%
  bind_cols(data.frame(H_donors = py_to_r(py$py_h_donors),
                       H_acceptors = py_to_r(py$py_h_acceptors),
                       TPSA = py_to_r(py$py_tpsa),
                       Rot_Bonds = py_to_r(py$py_rotbonds))
            ) %>%
  left_join(
    target_chem_pw_full
  ) %>%
  rowwise() %>%
  mutate(logD_7.4 = httk::calc_dow(dtxsid = dtxsid, pH = 7.4))

target_chem_lipinski <- target_chem_summary %>%
  select(-c(smiles)) %>%
  group_by(dtxsid) %>%
  mutate("Mass <= 500 Da" = (monoisotopicMass <= 500),
          "H-donors < 5" = (H_donors < 5),
          "H-acceptors < 10" = (H_acceptors < 10),
          "logP < 5" = (logkow < 5),
          .keep = "none") %>%
  ungroup() %>%
  mutate(Lipinski = rowSums(dplyr::select(., where(is.logical))) >= 3)

lipinski_plot <- target_chem_lipinski %>%
  group_by(Lipinski) %>%
  count(`Mass <= 500 Da`, `H-donors < 5`, `H-acceptors < 10`, `logP < 5`,
        name = "Count") %>%
  ggplot(aes(
    axis3 = `Mass <= 500 Da`,
    axis4 = `logP < 5`,
    axis2 = `H-donors < 5`,
    axis1 = `H-acceptors < 10`,
    y = Count
  )) +
  geom_alluvium(aes(fill = Lipinski)) +
  geom_stratum() +
  scale_x_discrete(limits = c( 
    "H-acceptors < 10",
    "H-donors < 5",
    "<= 500 Da",
    "logP < 5"
  ),
  position = "top",
  expand = c(.2, .05)) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            size = 2) +
  labs(title = "Target chemicals passing Lipinksi's Rules of 5",
       fill = "Passes Lipinksi's Rules?") +
  scale_fill_manual(values = c("grey20", "green4")) +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", color = "black", size = 10),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")

lipinski_plot
ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Alluvium_Lipinski.png"),
       bg = "white",
       height = 4, width = 5
)

chem_rell_label <- total_rel_evaluation_df %>% distinct(Chemical, ll_label) %>%
  filter(ll_label %in% c("Restrictive", "Nonrestrictive"))

# Molecular descriptors in bins
bin_size <- 10
logP_plot <- target_chem_summary %>%
  inner_join(chem_rell_label, by = join_by(dtxsid == Chemical)) %>%
  distinct(dtxsid, logkow) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(logP_bin = factor(logkow %/% bin_size * bin_size)) %>%
  ggplot(aes(x = logP_bin, y = Funbound.plasma)) +
  labs(x = "logKow", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

target_chem_summary %>%
  inner_join(chem_rell_label, by = join_by(dtxsid == Chemical)) %>%
  distinct(dtxsid, logkow, TPSA, Rot_Bonds, ll_label) %>% 
  pivot_longer(cols = c(logkow, TPSA),
               names_to = "Parameter", values_to = "Value") %>%
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  ggplot(aes(x = Value, y = Funbound.plasma)) +
  labs(x = "", y = "fup") +
  facet_grid(cols = vars(Parameter),
             rows = vars(ll_label),
             scales = "free_x") +
  geom_point() +
  theme_bw()

bin_size <- 50
TPSA_plot <- target_chem_summary %>%
  distinct(dtxsid, TPSA) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(TPSA_bin = factor(TPSA %/% bin_size * bin_size)) %>%
  ggplot(aes(x = TPSA_bin, y = Funbound.plasma)) +
  labs(x = "Topological Polar Surface Area", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 2
RotBond_plot <- target_chem_summary %>%
  distinct(dtxsid, Rot_Bonds) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(Rot_Bonds_bin = factor(Rot_Bonds %/% bin_size * bin_size)) %>%
  ggplot(aes(x = Rot_Bonds_bin, y = Funbound.plasma)) +
  labs(x = "Rotatable Bonds", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 150
MW_plot <- target_chem_summary %>%
  distinct(dtxsid, monoisotopicMass) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(MW_bin = factor(monoisotopicMass %/% bin_size * bin_size)) %>%
  ggplot(aes(x = MW_bin, y = Funbound.plasma)) +
  labs(x = "Molecular Mass", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 4
HBA_plot <- target_chem_summary %>%
  distinct(dtxsid, H_acceptors) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(HBA_bin = factor(H_acceptors %/% bin_size * bin_size)) %>%
  ggplot(aes(x = HBA_bin, y = Funbound.plasma)) +
  labs(x = "Hydrogen bond acceptors", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

>>>>

```{r data_loading}
this_pk <- readRDS("data-raw/plasma_CvTfits.rds")
load("data-raw/chem_lists.RData")
load("data-raw/plasma_CvTfits_data.RData")
load("data-raw/rat_pars_human.RData")
all_chem_properties <- readRDS("data-raw/ctxR_3comp2_cheminfo.rds")

# Sipes 2017 data
cyp_data <- readr::read_csv(paste0(
  Sys.getenv("FIG_DIR"),
  "RestrictiveClearance_Analysis/",
  "CLint_CYP_Sipes2017.csv"
)) %>%
  left_join(
    chem.physical_and_invitro.data %>%
      select(CAS = CAS, Chemical = DTXSID) %>%
      distinct()
  ) %>%
  select(Chemical, CAS, starts_with("MET_")) %>%
  filter(!is.na(Chemical))
names(cyp_data) <- c("Chemical", "CAS", "CYP1A2",
                     "CYP2C9", "CYP2C19", "CYP2D6", "CYP3A4")
# Functional use categories from CPdat
functional_uses_all <- readr::read_csv(paste0(
  Sys.getenv("FIG_DIR"), "RestrictiveClearance_Analysis/",
  "ChemExpo_bulk_functional_uses_20250304.csv"
))
```

@@

```{r venn_diagrams}

ggvenn(member_list, fill_color = c("white", "white"), auto_scale = FALSE)

@@

combined_RMSLE <- combined_preds %>%
  ungroup() %>%
  mutate(
    restrictive_SLE = (log1p(httk_Preds_restrictive) - log1p(Conc))^2,
    nonrestrictive_SLE = (log1p(httk_Preds_nonrestrictive) - log1p(Conc))^2,
    invivoPK_SLE =  (log1p(Conc_est) - log1p(Conc))^2 
  ) %>%
  group_by(Chemical, Species, Route) %>%
  summarize(RMSLE_restrictive = sqrt(mean(restrictive_SLE, na.rm = TRUE)),
            RMSLE_nonrestrictive = sqrt(mean(nonrestrictive_SLE, na.rm = TRUE)),
            RMSLE_invivoPKfit = sqrt(mean(invivoPK_SLE, na.rm = TRUE))) %>%
  ungroup() %>%
  mutate(comparison_group = case_when(
    between(RMSLE_invivoPKfit, RMSLE_nonrestrictive, RMSLE_restrictive) |
     between(RMSLE_invivoPKfit, RMSLE_restrictive, RMSLE_nonrestrictive) ~ "Between",
    RMSLE_invivoPKfit < RMSLE_nonrestrictive &
      RMSLE_invivoPKfit < RMSLE_restrictive ~ "Less than",
    RMSLE_invivoPKfit > RMSLE_restrictive &
      RMSLE_invivoPKfit > RMSLE_nonrestrictive~ "More than"
  ))

combined_RMSLE %>%
  ggplot(aes(x = Chemical, y = RMSLE_invivoPKfit)) +
  geom_point(color = "green4") +
  geom_point(aes(y = RMSLE_nonrestrictive), shape = 6, size = 0.5) +
  geom_point(aes(y = RMSLE_restrictive), shape = 4, size = 0.5) +
  geom_linerange(aes(ymin = RMSLE_nonrestrictive, ymax = RMSLE_restrictive)) +
  facet_grid(~factor(comparison_group,
                     levels = c("Less than", "Between", "More than")),
             scales = "free_x", space = "free_x") +
  labs(y = "RMSLE") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank(),
        axis.title = element_text(size = 14),
        strip.text = element_text(size = 12))

# AIC and relative likelihood
# LL from pre-calculated httk predictions
restrictive_ll <- all_httk_preds %>% 
  rename(Conc_est = httk_Preds_restrictive,
         Time_trans = Time) %>%
  select(!httk_Preds_nonrestrictive) %>%
  distinct() %>%
  group_by(Chemical, Species) %>%
  nest(.key = "observations") %>%
  inner_join(coef(this_pk) %>% semi_join(this_winmodel) %>% distinct()) %>%
  rowwise() %>%
  mutate(log_likelihood = log_likelihood(
    par = coefs_vector,
    data = observations,
    data_sigma_group = observations$data_sigma_group,
    dose_norm = FALSE,
    log10_trans = FALSE,
    negative = FALSE,
    includes_preds = TRUE,
    detect
  )) %>%
  select(!c(observations, coefs_vector, Time.Units, Time_trans.Units)) %>%
  mutate(model = "model_3comp2_restrictive",
         AIC = -2 * log_likelihood)

httk_AIC <- bind_rows(restrictive_ll,
                      nonrestrictive_ll,
                      this_AIC %>% filter(model == "model_flat") %>%
                        filter_targets())

this_rel_evaluated_df <- this_css_df %>%
  inner_join(httk_rel_ll %>% distinct(Chemical, Species, ll_label)) %>%
  inner_join(this_tkstats) %>% # Css
  inner_join(this_httk_clearance_df) %>% # CLtot and halflife
  inner_join(data.frame(this_clint_df)) %>%
  inner_join(this_param_df %>% 
               distinct(Chemical,
                        forced_human_values,
                        Fabsgut.restrictive,
                        Fabsgut.nonrestrictive)) %>%
  mutate(Chemical = forcats::fct_reorder(Chemical, restrictive_halflife)) %>%
  ungroup() %>%
  distinct()

halflife_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, Css_httk_r,
           halflife.tkstats, restrictive_halflife,
           nonrestrictive_halflife,
           ll_label) %>%
  ggplot(aes(x = Chemical)) +
  geom_pointrange(aes(y = halflife.tkstats, 
                      ymin = nonrestrictive_halflife,
                      ymax = restrictive_halflife),
                  size = 0.1, shape = 23) +
  facet_grid(cols = vars(ll_label), space = "free_x", scales = "free_x") +
  scale_y_log10() +
  labs(x = "Chemicals", y = "halflife (hours)") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank(),
        axis.text.y = element_text(color = "black", size = 10),
        axis.title = element_text(size = 10))

ggsave(halflife_pl, filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "halflife_fullChart.png"),
       bg = "white",
       height = 3, width = 12
)

css_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, Css.tkstats, Css_httk_nr, Css_httk_r,
         ll_label) %>%
  ggplot(aes(x = Chemical)) +
  geom_pointrange(aes(y = Css.tkstats, ymin = Css_httk_nr, ymax = Css_httk_r),
                  size = 0.1, shape = 23) +
  facet_grid(cols = vars(ll_label), space = "free_x", scales = "free_x") +
  scale_y_log10() +
  labs(x = 'Chemicals', y = "Css") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank(),
        axis.text.y = element_text(color = "black", size = 10),
        axis.title = element_text(size = 10))

cltot_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, CLtot_Fgutabs.tkstats,
           restrictive_clearance, nonrestrictive_clearance, 
           Fabsgut.restrictive, Fabsgut.nonrestrictive,
         ll_label) %>%
  ggplot(aes(x = Chemical)) +
  geom_pointrange(aes(y = CLtot_Fgutabs.tkstats,
                      ymin = restrictive_clearance/Fabsgut.restrictive,
                      ymax = nonrestrictive_clearance/Fabsgut.nonrestrictive),
                  size = 0.1, shape = 23) +
  facet_grid(cols = vars(ll_label), space = "free_x", scales = "free_x") +
  scale_y_log10() +
  labs(x = 'Chemicals', y = "CLtot/Fabsgut") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank(),
        axis.text.y = element_text(color = "black", size = 10),
        axis.title = element_text(size = 10))

(halflife_pl + theme(axis.title.x = element_blank())) /
  (css_pl+ theme(axis.title.x = element_blank())) /
  cltot_pl

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "tkcomparisons_fullChart.png"),
       bg = "white",
       height = 4.5, width = 12
)

This is the enrichment analysis for functional use categories.

```{r functional}
all_chem_functions <- functional_uses_all %>%
  filter(DTXSID %in% all_chems, !is.na(`Harmonized Functional Use`)) %>% 
  group_by(`Harmonized Functional Use`) %>% 
  count() %>% arrange(-n) %>% head(20) %>% pull(1)
target_chem_functions <- functional_uses_all %>%
  filter(DTXSID %in% target_chems, !is.na(`Harmonized Functional Use`)) %>% 
  group_by(`Harmonized Functional Use`) %>% 
  count() %>% arrange(-n) %>% head(20) %>% pull(1)

combined_function_list <- union(all_chem_functions,
                                target_chem_functions) # 25 total

pre_enrichment_mat <- inner_join(
  functional_uses_all %>%
    filter(
      DTXSID %in% all_chems, !(DTXSID %in% target_chems), # all non-targets
      `Harmonized Functional Use` %in% combined_function_list
    ) %>% distinct(DTXSID, `Harmonized Functional Use`) %>%
    count(`Harmonized Functional Use`, name = "n_nontargets"),
  functional_uses_all %>%
    filter(
      DTXSID %in% target_chems, 
      `Harmonized Functional Use` %in% combined_function_list
    ) %>% distinct(DTXSID, `Harmonized Functional Use`) %>%
    count(`Harmonized Functional Use`, name = "n_targets")
) %>%
  janitor::clean_names() %>%
  mutate(n_total_targets = length(target_chems),
         n_total_nontargets = length(all_chems) - n_total_targets,
         n_nontargets_out = n_total_nontargets - n_nontargets,
         n_targets_out = n_total_targets - n_targets) %>%
  select(!c(n_total_nontargets, n_total_targets))

quick_enrich <- pre_enrichment_mat %>%
  rowwise() %>%
  mutate(odds_ratio = fisher.test(
    matrix(c(n_targets, n_nontargets,
             n_targets_out, n_nontargets_out),
           nrow = 2),
    alternative = "greater")$estimate,
    pval = fisher.test(
    matrix(c(n_targets, n_nontargets,
             n_targets_out, n_nontargets_out),
           nrow = 2),
    alternative = "greater")$p.value
  ) %>%
  ungroup() %>%
  mutate(padj = 0.05/n()) %>%
  arrange(pval)

# Makeshift upset plot
upset_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, Css_httk_r,
           halflife.tkstats, restrictive_halflife,
           nonrestrictive_halflife,
           ll_label) %>%
left_join(functional_uses_all %>%
            distinct(DTXSID, `Harmonized Functional Use`) %>%
            rename(Chemical = DTXSID) %>%
            filter(`Harmonized Functional Use` %in% c(
              "Biocide",
              "Pharmaceutical (EPA)",
              "Degradant/impurity (EPA)"))
          ) %>%
  # filter(!is.na(`Harmonized Functional Use`)) %>%
  mutate(val = ifelse(is.na(`Harmonized Functional Use`), NA, "Present")) %>%
  distinct(Chemical, ll_label, `Harmonized Functional Use`, val) %>%
  ggplot(aes(x = Chemical, y = `Harmonized Functional Use`, color = val)) + 
  geom_point() +
  scale_color_manual(values = "black") +
  facet_grid(cols = vars(ll_label), scales = "free_x", space = "free_x") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(color = "white"),
        strip.text = element_blank(),
        legend.title = element_blank(), axis.title.y = element_blank())

(halflife_pl + theme(axis.title = element_blank())) /
  (css_pl + theme(axis.title = element_blank())) /
  upset_pl

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "tkcomparisons_withUpset.png"),
       bg = "white",
       height = 4.5, width = 12
)

quick_enrich$padj_val <- p.adjust(quick_enrich$pval, "bonferroni")

# Enrichment plot
quick_enrich %>%
  mutate(
    harmonized_functional_use = forcats::fct_reorder(
      harmonized_functional_use, odds_ratio)
    ) %>%
  filter(padj_val <= 0.5) %>%
  ggplot() +
  geom_col(aes(x = odds_ratio, 
               y = harmonized_functional_use,
                fill = padj_val),
           color = "black", linewidth = 0.2,
           width = 0.75) +
  scale_fill_viridis_b(option = "viridis", direction = -1,
                       breaks = c(0, 0.05, 0.1, 0.5),
                       limits = c(0, 0.5)
                       ) +
  scale_x_continuous(limits = c(0, 12),
                     expand = expansion()) +
  labs(x = "Enrichment Score", y = "",
       title = "Enrichment of Functional Use Categories",
       fill = "adjusted p-value") +
  theme_bw() +
  theme(legend.title = element_text(hjust = 0.5),
        axis.text = element_text(color = "black"),
        axis.text.y = element_text(face = "bold", size = 10),
        plot.title = element_text(hjust = 0.5))

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "FunctionalUse_Enrichment.png"),
       bg = "white",
       height = 5, width = 6.5
)

@@

```{r selecting_chemical_params}

# Testing to make sure there are some chemicals that have different clearances
count(this_httk_clearance_df, restrictive_clearance == nonrestrictive_clearance)
# The majority (140) do!

```{r load-other}

# Check to see how much of the full dataset is available
length(intersect(unique(new.raw.httk.data$DTXSID), full_list))

focus_chems <- union(new.raw.httk.data$DTXSID, target_chems)

all_chem_properties %>%
  filter(propertyId %in% c("logkow-octanol-water",
                     "pka-acidic-apparent",
                     "pka-basic-apparent",
                     "water-solubility",
                     "logkoa-octanol-air",
                     "henrys-law",
                     "boiling-point")) %>%
  glimpse()

@@

all_hum_rat_fups <- chem.physical_and_invitro.data %>%
  filter(DTXSID %in% all_chems) %>%
  select(DTXSID, Human.Funbound.plasma, Rat.Funbound.plasma) %>%
  mutate(Rat.Funbound.plasma = as.numeric(str_extract(Rat.Funbound.plasma,
                                                      "^([^,]+)")),
         Human.Funbound.plasma = as.numeric(str_extract(Human.Funbound.plasma,
                                                        "^([^,]+)"))
         )

# Get restrictive and non-restrictive parameters

@@

kd_v_fup_plot <- new.raw.httk.data %>%
  inner_join(all_hum_rat_fups) %>%
  pivot_longer(cols = Human.Funbound.plasma:Rat.Funbound.plasma,
               names_to = "Species", values_to = "httk_fup_med") %>%
  ggplot(aes(x = Affinity.Kd.Med, y = httk_fup_med, color = Species)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("green4", "purple2"),
                     labels = c("Human", "Rat")) +
  scale_x_log10(labels = scales::label_log(digits = 2)) +
  labs(x = "Dissociation constant Kd\n(uM, from in vitro RED assay)",
       y = "Fraction unbound\nin plasma") +
  theme_bw() +
  theme(legend.position = "left",
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8, color = "black"))

kd_v_fup_plot

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Kd_estimates_vs_fup_multispecies.png"),
       bg = "white",
       height = 2.8, width = 5
)

@@
